Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-16 csf-2017/paper.tex
figure 3 and scall

\begin{figure}
  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
    \draw[draw=gray!80] (0,0.5) -- (0,4);
    \draw[draw=gray!80] (6,0.5) -- (6,4);
    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous program counter};
    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous stack pointer};
    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
    \draw (-3,2.25) node (rp) {return pointer};
    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,2.25);
  \end{tikzpicture}

@@

\textbf{Calling convention} With the security measures introduced and motivated,
here is a summary of our proposed calling convention:
\begin{description}[font=\normalfont\itshape]
\item[At program start-up] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure the callback capability is global.
\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} We define a number of reusable macros
capturing the calling convention and other conveniences. All macros that uses
the stack assumes a stack pointer in register $r_\stk$.
\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
$\var{name}$ from the linking table and stores it in register $r$.
\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
invoked by the adversary and prepares the received stack pointer by checking
that it has permission $\rwlx$.
\texttt{\footnotesize{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
capability in register $r$ in the manner described above. That is, it pushes
local state (the contents of registers $\overline{r_{\var{priv}}}$), return
pointer and stack pointer to the stack, sets up the necessary on stack return
code and creates an $\entry$ return pointer, restricts the stack pointer, clears
the unused part of the stack, clears the necessary registers and jumps to $r$.
Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
clears all the memory the capability in register $r$ has authority over.
\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
register $r$ is a global capability. \texttt{\footnotesize{crtcls
    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
closure's code and a new environment is allocated (using malloc) where the
contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

>>>>

\begin{figure}
  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
    \draw[draw=gray!80] (0,0.5) -- (0,4);
    \draw[draw=gray!80] (6,0.5) -- (6,4);
    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous stack pointer};
    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous program counter};
    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
    \draw (-3,3.25) node (rp) {return pointer};
    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,3.25);
  \end{tikzpicture}

@@

\textbf{Calling convention} With the security measures introduced and motivated,
here is a summary of our proposed calling convention:
\begin{description}[font=\normalfont\itshape]
\item[At program start-up] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure the callback capability is global.
\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} We define a number of reusable
macros capturing the calling convention and other conveniences. All
macros that uses the stack assumes a stack pointer in register
$r_\stk$.  \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
capability related to $\var{name}$ from the linking table and stores
it in register $r$.  \texttt{\footnotesize{push $r$}} and
\texttt{\footnotesize{pop $r$}} work as one would
expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
invoked by the adversary and prepares the received stack pointer by
checking that it has permission $\rwlx$.  \texttt{\footnotesize{scall
    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}}
jumps to the capability in register $r$ in the manner described
above. That is, it pushes local state (the contents of registers
$\overline{r_{\var{priv}}}$), the necessary on stack return code,
return pointer, stack pointer to the stack, and creates an $\entry$
return pointer, restricts the stack pointer, clears the unused part of
the stack, clears the necessary registers and jumps to $r$.  Upon
return, the private state is restored. \texttt{\footnotesize{mclear
    $r$}} clears all the memory the capability in register $r$ has
authority over.  \texttt{\footnotesize{rclear $\var{regSet}$}} clears
all the registers in $\var{regSet}$. \texttt{\footnotesize{reqglob
    $r$}} checks whether the word in register $r$ is a global
capability. \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}}
allocates a closure where $r$ points to the closure's code and a new
environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the code referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.
